---
title: "class08: Halloween Candy Mini-Project"
author: "Jiachen Fan (A17662703)"
format: pdf
---
# 1. Importing candy data
First things first, let’s get the data from the FiveThirtyEight GitHub repo.
```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```
> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```
85 different candy types.

> Q2. How many fruity candy types are in the dataset?

```{r}
table(candy$fruity)
```
38 fruity candy types.

# 2. What is your favorate candy?

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?


```{r}
candy["Fruit Chews", ]$winpercent
```
My favorite candy is 'Fruit Chews' and its winpercent value is 43.08892.

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
# install.packages('skimr')
library("skimr")
skim(candy)
```
> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Winpercent. Others are on a 0 to 1 scale.

> Q7. What do you think a zero and one represent for the candy$chocolate column?

0 means there is no chocolate in ihe candy. 1 means there is chocolate.

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```
> Q9. Is the distribution of winpercent values symmetrical?

No.

> Q10. Is the center of the distribution above or below 50%?

Below 50%.

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean(candy$winpercent[as.logical(candy$chocolate)]) > mean(candy$winpercent[as.logical(candy$fruity)])
```

Yes.

> Q12. Is this difference statistically significant?

```{r}
t.test(candy$winpercent[as.logical(candy$chocolate)], candy$winpercent[as.logical(candy$fruity)])
```

Yes it is.

# 3. Overall Candy Rankings
> Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n = 5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n = 5)
```

> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, rownames(candy))+
  geom_col()

```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy),winpercent))+
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```
```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
> Q17. What is the worst ranked chocolate candy?

It is 'Sixlets'.

> Q18. What is the best ranked fruity candy?

It is 'Starburst'.

# 4. Taking a look at pricepercent

```{r}
# install.packages('ggrepel')
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 100)  
# change 'max.overlaps' to 100 after warning
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

It is Reese's Miniatures.

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

Above are the top 5 most expensive candy types and 'Nik L Nip' is the least popular.

> Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

```{r}
# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
  geom_point()
```

# 5 Exploring the correlation structure

```{r}
# install.packages('corrplot')
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```
> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?


Fruity and chocolate.

> Q23. Similarly, what two variables are most positively correlated?

Winpercent and chocolate.

# 6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2])
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
```{r}
# Make a new data-frame 
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
      aes(x=PC1, y=PC2,size=winpercent/100,text=rownames(my_data), label=rownames(my_data)) +
      geom_point(col=my_cols)

p
```
Again we can use the ggrepel package and the function ggrepel::geom_text_repel() to label up the plot with non overlapping candy names like. 

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 100)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
```{r}
library(plotly)
```

```{r}
#| eval: false

ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard and pluribus are picked up strongly by PC1 in the positive direction.
Yes, I think fruity candy types are usually hard and packaged in bags or boxes. These variables are positively correlated.