---
title: "class05: Data Visualization with ggplot"
author: "Jiachen Fan (A17662703)"
format: pdf
---

## Use ggplot

To use ggplot we first need to install it. Use 'install.packages()'.
Before I use any package functions I have to load them up with a 'library()'.

```{r}
library(ggplot2)

ggplot(cars)

```
```{r}
tail(cars,6)
```
```{r}
ggplot(cars) +
  aes(x=speed, y=dist)
```
```{r}
ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point()
```

To use ggplot, three things matter
-data(data.frame)
-aesthetics (how the data map to the plot)
-geoms (how I want things drawn)

```{r}
ggplot(cars)+
  aes(x=speed, y=dist)+
  labs(title ='Speed and Stopping Distance of the Cars',
       x = 'Speed(MPH)',
       y = 'Stopping distance(ft)') +
  geom_point()+
  geom_smooth(method ='lm') +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
```
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
```{r}
nrow(genes)
colnames(genes)
ncol(genes)
table(genes$State)
round(table(genes$State)/nrow(genes)*100,2)
```
```{r}
ggplot(genes)+
  aes(x= Condition1, y=Condition2) +
  geom_point()
```
```{r}
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point()
p
```
```{r}
p + scale_colour_manual( values=c("blue","gray","red") )+
  theme_bw()
```
```{r}
p + scale_colour_manual( values=c("blue","gray","red") )+
  labs(title = 'Gene Expression Changes Upon Drug Treatment',
       x = 'Control (no drug)',
       y = 'Drug Treatment') +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
```
# 7.Going Further
```{r}
library(gapminder)
```
```{r}
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)

# install.packages("dplyr")  ## un-comment to install if needed
library(dplyr)
gapminder_2007 <- gapminder %>% filter(year==2007)
```
```{r}
ggplot(gapminder_2007)+
  aes(x = gdpPercap, y = lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.4)
```
```{r}
ggplot(gapminder_2007)+
  aes(x = gdpPercap, y = lifeExp, color=pop) +
  geom_point(alpha=0.4)
```
```{r}
q <- ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, size = pop) +
  geom_point(alpha=0.4)
q
```
```{r}
q +  scale_size_area(max_size = 10)
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)
ggplot(gapminder_1957)+
  aes(x = gdpPercap, y = lifeExp, color = continent, size = pop) +
  geom_point(alpha = 0.4) +
  scale_size_area(max_size = 10)
```
```{r}
gapminder_1957_2007 <- gapminder %>% filter(year==1957|year==2007)
s <-ggplot(gapminder_1957_2007)+
  aes(x = gdpPercap, y = lifeExp, color = continent, size = pop) +
  geom_point(alpha = 0.4) +
  scale_size_area(max_size = 10) +
  facet_wrap(~year)
s
```

```{r}
table(gapminder$continent)
s + scale_colour_manual( values=c("blue","bisque2",
                                  "darkslategray2","chartreuse1","darkolivegreen3") ) +
  theme_bw()
```

# 8.Bar Charts
```{r}
gapminder_top5 <- gapminder %>% 
  filter(year==2007) %>% 
  arrange(desc(pop)) %>% 
  top_n(5, pop)

gapminder_top5
```
```{r}
ggplot(gapminder_top5) + 
  aes(x = country, y = pop) +
  geom_col()
```
```{r}
ggplot(gapminder_top5) + 
  aes(x = country, y = lifeExp, fill = continent) +
  geom_col()
```
```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop, fill = lifeExp))
```
```{r}
ggplot(gapminder_top5) +
  aes(x=country, y=pop, fill=gdpPercap) +
  geom_col()
```
```{r}
ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=gdpPercap) +
  geom_col()
```
```{r}
ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=country) +
  geom_col(col="gray30") +
  guides(fill="none")
```
```{r}
head(USArrests)
```
```{r}
USArrests$State <- rownames(USArrests)
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_col() +
  coord_flip()
```
```{r}
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_point() +
  geom_segment(aes(x=State, 
                   xend=State, 
                   y=0, 
                   yend=Murder), color="blue") +
  coord_flip()
```
# 9.Extensions: Animation
First, install.packages("gifski") and install.packages("gganimate")
Second, use '#| eval:false' to skip the execution of animation.
```{r}
#| eval: false
library(gapminder)
library(gganimate)

ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # gganimate 
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  shadow_wake(wake_length = 0.1, alpha = FALSE)
```
# 10.Combining plots

Install 'patchwork'
```{r}
library(patchwork)

p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))
p3 <- ggplot(mtcars) + geom_smooth(aes(disp, qsec))
p4 <- ggplot(mtcars) + geom_bar(aes(carb))

(p1 | p2 | p3) /
      p4
```

