---
title: "homework: R functions"
author: "Jiachen Fan (A17662703)"
format: pdf
---
# A. Improve this regular R code by abstracting the main activities in your own new function. Note, we will go through this example together in the formal lecture. The main steps should entail running through the code to see if it works, simplifying to a core working code snippet, reducing any calculation duplication, and finally transferring your new streamlined code into a more useful function for you.
Let's have a look at original code
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
df
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a)) # normalize values in df$a 
df
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
df
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c)) # normalize values in df$c
df
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))
df
```
Now we can simplify it using function 'normalize'
```{r}
# Define a normalization function
normalize <- function(x) {
    (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
}

# Create data frame
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)

# Apply normalize to df$a and df$c
df[c('a','c')] <- apply(df[c('a','c')], 2, normalize)
df$b <- normalize(df$b)+1         
#(df$b - min(df$a)) / (max(df$b) - min(df$b)) = (df$b - min(df$b)+min(df$b)-min(df$a)) / (max(df$b) - min(df$b)) = normalize(df$b) + 1/1
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))

# It is hard to improve because the codes are different for each column. It will be much easier if the original code meant to normalize all the columns.

```
# B. Next improve the below example code for the analysis of protein drug interactions by abstracting the main activities in your own new function. Then answer questions 1 to 6 below. It is recommended that you start a new Project in RStudio in a new directory and then install the bio3d package noted in the R code below (N.B. you can use the command install.packages("bio3d") or the RStudio interface to do this). Then run through the code to see if it works, fix any copy/paste errors before simplifying to a core working code snippet, reducing any calculation duplication, and finally transferring it into a more useful function for you.
