---
title: "class11: Comparative structure analysis (Pt2)"
author: "Jiachen Fan (A17662703)"
format: pdf
---
# 4. Comparative structure analysis of Adenylate Kinase

Install packages in the R console NOT your Rmd/Quarto file

install.packages("bio3d")
install.packages("devtools")
install.packages("BiocManager")

We use 'BiocManager::install("msa")' to install any other BioConductor package we want
devtools::install_bitbucket("Grantlab/bio3d-view")

```{r}
library(bio3d)
```


> Q10. Which of the packages above is found only on BioConductor and not CRAN?

msa

> Q11. Which of the above packages is not found on BioConductor or CRAN?


> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?

True


```{r}
aa <- get.seq('1ake_A')
aa
```


> Q13. How many amino acids are in this sequence, i.e. how long is this sequence?

214
```{r}
#b <- blast.pdb(aa)
```

```{r}
# Plot a summary of search results
#hits <- plot(b)
```
```{r}
# List out some 'top hits'
#hits$pdb.id
```
> side-note: Lets save our results thus far so we don't have to run blast again.

```{r}
#save(hits, b, file='blast_results.Rds')
```

```{r}
load('blast_results.Rds')
hits
```

Download all these related structures from the database with 'get.pdb()'
```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
# Align and superpose structures
```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```
# Principal component analysis

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```
```{r}
dim(pdbs$xyz)
```
```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

```{r}
mktrj(pc.xray, file = 'pca_results.pdb')
```

